### 2017-02-16 real/draft.tex
[draft] Polish "numerical implementation"
Remove several comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Bayesian prediction of luminosity distribution based on one simulation run:
% An example from TARDIS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}

\usepackage[a4paper, left=20mm, right=20mm, top=25mm, bottom=25mm]{geometry}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage{physics}

\usepackage[colorinlistoftodos]{todonotes} % Use 'disable' to remove todos in final version
\newcommand{\fred}[1]{\todo[color=orange!40,inline]{#1}} %
\newcommand{\fredmargin}[1]{\todo[color=orange!40]{#1}} %
\newcommand{\checked}{\todo[color=green,noline]{\checkmark}} %
\newcommand{\checkedl}{\todo[color=brown]{\checkmark}} %
\newcommand{\hans}[1]{\todo[color=yellow!30,inline]{#1}} %
\newcommand{\hanslong}[2][]{\todo[%bordercolor=red,
  color=white,inline,caption={2do}, #1]{
    \begin{minipage}{\textwidth} #2\end{minipage}}} %

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{170212}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}

\parindent=3ex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm,amssymb,amsfonts,amsmath}  % bold math symbols, AMStex
\usepackage{graphicx}    % standard graphics
\graphicspath{
  {./figures/}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\lleq}[1]{\label{#1} }
% TO REMOVE THE LABEL LETTERS FROM THE EQUATION DISPLAY, COMMENT OUT THIS LINE:
\renewcommand{\lleq}[1]{\label{#1} {\scriptstyle {\rm (#1)}} \hspace*{2ex} }

\input{symbols}

\newcommand{\hypo}  {{\mathcal{H}}}  % Model
\newcommand{\ldef}{\;{:}{=}\;}
\newcommand{\rdef}{\;{=}{:}\;}
\newcommand{\realnumbers}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\cond}{\,|\,}
\newcommand{\NBD}{\mathrm{NB}}

\newcommand{\refeq}[1]{Eq.~(\ref{#1})}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\refsec}[1]{Sec.~\ref{sec:#1}}
\DeclareMathOperator{\erfc}{erfc}
\DeclareMathOperator{\Expect}{\mathbb{E}}
\newcommand{\expect}[1]{\Expect\left[#1\right]}
\newcommand{\expectest}[1]{\widehat{\Expect\left[#1\right]}}
\DeclareMathOperator{\GammaDist}{Gamma}
\DeclareMathOperator{\GaussianDist}{\mathcal{N}}
\DeclareMathOperator{\InvGammaDist}{InvGamma}
\newcommand{\Kalpha}{{K_\alpha}}
\newcommand{\Kbeta}{{K_\beta}}
\newcommand{\npack}{{n_p}}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{{L_{\rm max}}}
\newcommand{\Lumtot}{Q}
\newcommand{\lumtot}{q}
\newcommand{\Lum}{L}
\newcommand{\lum}{\ell}
% \newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\rmdx}[1]{\dd{#1}} % differential
\newcommand{\firstDeriv}[1]{\frac{\partial}{\partial #1}}
\newcommand{\secDeriv}[1]{\frac{\partial^2}{\partial #1^2}} % second partial derivative
\newcommand{\secPartial}[2]{\frac{\partial^2}{\partial #1 \, \partial #2}} % second partial derivative
\newcommand{\tardis}{TARDIS}
\DeclareMathOperator{\Variance}{\mathbb{V}}
\newcommand{\variance}[1]{\Variance\left[#1\right]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{center}
  \textbf{\Large Bayesian prediction of supernova luminosity  distributions}\\[8pt]
  \textbf{\Large based on simulation output}\\[12pt]
\end{center}

\textbf{Abstract} TODO\\

\section{What's left to do}

\subsection{why asymptotic approx. has larger variance}
\begin{enumerate}
  \item For $\alpha=1.5, \beta=60, n=14000$ still get 50 \% larger uncertainty from asymptotic compared to summation over $N$. Is it due very asymmetric one-packet distribution? Try larger $\alpha$, for example $\alpha = 30$.
  \item  Another option: $\int \rmdx{\alpha} \rmdx{\beta}$, use true values in simulation. Where does uncertainty come from? From $\lambda$ or $\mu, \sigma^2$?
  \item Are finite number of terms in $\sum_N$ a problem? Check

\end{enumerate}

\subsection{Plots}

\begin{enumerate}
  \item generate a toy spectrum, take frequencies from actual tardis,
  generate luminosities from a $\GammaDist(\alpha, \beta)$, no $\nu$
  dependence of $\alpha, \beta$ for clarity , plot spectrum with
  uncertainty bands, show it early in the paper, then describe where
  it comes from
  \item compare asymptotic and $\sum_N$ $p(\Lumtot \cond \bmy)$ on a
  plot, add what happens if only $\sqrt{n}$ uncertainty is used: that
  would neglect uncertainty on the single-packet distribution, show
  case when parameters determined from a single bin with few samples,
  should make large difference. Fix $\lambda$, what are the uncertainties now?
  \item Statistics of replicas: again compare \refeq{spe} to \refeq{ral}. Teach a lesson: from a single data set, prediction will be centered there
  \item Question: does Wolfgang want to see a plot in which ``$\lambda$ is divided out''? That is , just plot how mean
\end{enumerate}

\subsection{Other issues that need to be resolved}

\begin{enumerate}

\item Clean up integrals using \verb=\rmdx= rather than $d$.

\item Write everything in terms of frequency $\nu$ or wavelength
  $\lambda$

\item Now that we have removed direct analysis of \tardis\ from the
  paper, there is no need to define an overall parameterization of the
  Gamma parameters $\alpha(\nu_b) = \sum_k a_k \nu_b^k$ and
  $\beta(\nu_b) = \sum_k b_k \nu_b^k$. However, the possibility of
  linking the analyses of all the bins by this and similar methods
  should be highlighted. Perhaps in the Conclusions?

\item \textbf{\tardis} move tardis section into separate document in
  this git repo. Better to keep it out of the paper. Cite tardis when
  showing the toy spectrum. Mention that in reality $\alpha$ and
  $\beta$ are $\nu$ dependent

\end{enumerate}

\section{Introduction (Wolfgang)}

\subsection{Astrophysical context}

Wolfgang's text

\subsection*{Main points of this paper}

\begin{enumerate}
\item The first goal of this paper is to show that the Bayesian
  approach to this problem is well-founded and robust.

\item Luminosity data generated by a single simulation provide a
  basis for predicting its uncertainty, i.e.\ how much this luminosity
  can be expected to vary from simulation to simulation.

\item The uncertainty in the luminosity calculated in the Bayesian
  approach is larger than anticipated because it takes into account
  not only the variability of the luminosity per packet, but also the
  distribution of the number of packets.

\item Traditional methods using conventional counting ratios
  (``relative frequencies'') can also make predictions for such
  uncertainties, but only if the number of packets is large. By
  contrast, the Bayesian method presented here remains valid also for
  small packet multiplicities.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bayesian prediction}
\label{sec:bayp}

\fred{This is a quick summary of Bayesian inference, we could refer to some text books for more details}
We first sketch the generic notation and relations for Bayesian
probability theory.
%
A particular experiment or simulation generates
\textit{data}
which will often come in the form of a list of $n$ measurements,
$\bmy = (y_1,y_2,\ldots,y_n)$. In order to understand or explain these
data, the observer constructs one or more \textit{models} or
\textit{hypotheses} $\hypo$. Central to this construction is the
specification of a \textit{likelihood} $p(\bmy\cond \bmtheta,\hypo)$
where $\bmtheta = (\theta_1,\theta_2,\ldots,\theta_K)$ is the set of
unknown parameters specified by the model. The likelihood is the
probability assigned by the model $\hypo$ to the data $\bmy$ when the
parameters $\bmtheta$ are fixed and known.
%
If and when the individual measurements $y_i$ are independent, the
joint likelihood is the product
\begin{align}
  \lleq{bta}
  p(\bmy\cond \bmtheta) &= \prod_{i=1}^n p(y_i\cond \bmtheta).
\end{align}
At this level of analysis, the goal is to find which values for these
parameters provide a satisfactory description of the data. In the
Bayesian context, these parameter values are best described in terms
of the probability of the parameters given the data,
$p(\bmtheta\cond \bmy)$. This so-called \textit{posterior} is related
to the likelihood by Bayes' Theorem,
\begin{align}
  \lleq{bth}
  p(\bmtheta\cond \bmy)
  &= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    {p(\bmy)}
    %= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    %{\int \rmdx{\bmtheta}'\, p(\bmy\cond\bmtheta')\;p(\bmtheta')}
\end{align}
where $p(\bmtheta)$, the so-called \textit{prior}, is the probability
assigned to the values of the parameters before any data were taken
into account and the \textit{evidence}
\begin{align}
  \lleq{bti}
  p(\bmy) &= \int \rmdx{\bmtheta}'\, p(\bmy\cond\bmtheta')\;p(\bmtheta'),
\end{align}
is fully determined by the likelihood and prior.
%
Prediction of any ``future'' data $Y$ is made not by considering the
likelihood $p(Y\cond\bmtheta_{\rm max})$ using single
maximum-likelihood parameter values $\bmtheta_{\rm max}$, but by
integrating over all possible values of the likelihood, weighted by
the posterior,
\begin{align}
  \lleq{btj}
  p(Y\cond \bmy) &= \int \rmdx{\bmtheta}\,p(Y\cond\bmtheta)\,p(\bmtheta\cond \bmy).
\end{align}
The above relations are true for any type of data and can be applied
to actual experimental data and the pertinent physical parameters. The
scope of this paper, however, is limited to the relationship between
simulated data generated by radiative transfer simulation codes.
Likewise, all parameters considered below relate strictly to model
descriptions of the simulation.
\fred{I find this paragraph confusing. Which relationship?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single-simulation data and multi-simulation prediction}

% \subsection{Output data from radiative transfer simulations}
% \subsection{Simulation data}

We consider a single computer simulation which generates a luminosity
spectrum from $\npack$ photon packets, each having a particular
frequency $\nu_i$ and luminosity $\lum_i$. Together, these frequencies
and luminosities form the raw simulation data
$\bmy = \{ (\nu_i,\lum_i): i = 1,\ldots,\npack\}$.  We divide up the
frequency spectrum into bins of width $\Delta\nu$ centered on
frequencies $\nu_b$ resulting in bin intervals
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$. \fred{The binning may be dictated by experimental data} Let $n_b$ be the number
of packets having a frequency falling into a particular bin $b$, and
let the vector $\bml_b = \{(\ell_{b,j}): j = 1,\ldots,n_b\}$ be the
set of luminosities of those packets. The data set is correspondingly
cast into the more compact histogram form
\begin{align}
  \lleq{sdb}
  \bmy = \{(n_b,\bml_b): b = 1,\ldots,B\}
\end{align}
with loss of information limited to replacing the exact packet
frequencies $\nu_i$ by the approximate bin frequency $\nu_b$. The
total luminosity in bin $b$ is
\begin{align}
  \lleq{sdc}
  \lumtot_b
  &= \sum_{j=1}^{n_b} \lum_{b,j}.
\end{align}
Within a given model $\hypo$, the data $\bmy$ are only one of many
possible realisations of the corresponding random variables; for
example, starting the simulation with a different seed would result in
different output data. We denote the general random variables by the
corresponding upper-case symbols,
\begin{align}
  \lleq{sdd}
  \bmL_b &= \{(\Lum_{b,j}): j = 1,\ldots,N_b\}\\
  \lleq{sde}
  \bmY &= \{(N_b, \bmL_b): b = 1,\ldots,B\},\\
  \lleq{sdf}
  \Lumtot_b &= \sum_{j=1}^{N_b} \Lum_{b,j},
\end{align}
Lower-case symbols such as $\bmy, n_b,\bml_b$ etc.\ are strictly
reserved for the simulated data at hand.

The goal of this paper is to illustrate how the single-simulation data
$\bmy$ can be used to predict the distribution of total bin
luminosities $Q_b, b = 1,\ldots,B$ if many simulations had been
carried out. At this introductory level, we assume that the
luminosities at different frequencies are independent, so that the
prediction of the entire spectrum (including all its uncertainty) is
the product of predictions for each frequency bin,
\begin{align}
  \lleq{sdg}
  p(Q_1,Q_2,\ldots,Q_B\cond\bmy) = \prod_{b=1}^B p(Q_b\cond \bmy).
\end{align}

\section{The generic model} \label{sec:model}

Various physical processes such as absorption result in luminosity
spectra with strong features. The model we develop below takes into
account the variability introduced by these features in a two-step
process termed a \textit{compound process} in the statistics
literature.
%
In the first step, the number of packets $N$ is modelled by means of a
separate Poisson distribution $p(N\cond\lambda)$ for each bin, where
the magnitude of $\lambda$ will capture most of the variability of the
spectrum. Given $N$, the single-packet luminosity distribution is
modelled in the second step by means of a packet likelihood for
$\Lum_{b,j}$ with packet parameters, a packet prior and, of course,
packet parameter posterior. The single-packet distribution contains
the crucial information which allows us to extend previous methods to
cases where the number of packets in a bin is small.
%%%%%%%%%%%%%%%%%%%
\fred{Refs: Zech, maybe others, too}

We derive in this section the generic equations of the compound
process, and then proceed to show in Sections \ref{sec:example} and
\ref{sec:asymptotic} how these equations can be used in two different
ways.
%
We consider for the moment quantities in one bin $b$ only, and hence
drop the $b$-subscripts, e.g.\
$Q_b \to Q, n_b \to n, N_b\to N, \bml_b\to\bml, L_{b,j} \to L_j$ etc.

We seek a Bayesian prediction for the total luminosity in one bin,
$p(Q\cond n,\bml)$, one of the factors in Eq.~(\ref{sdg}).  To do so,
we must take into account all possible values of variables and
parameters entering our model.  This involves multiple use of the
\textit{marginalisation rule}, which states that, to obtain the
probability for a variable $U$ when we are not interested in another
variable $V$, the marginal probability for $U$ is %
$p(U) = \int dV\,p(U,V) = \int dV\,p(U\cond V)\,p(V)$ for continuous
$V$ and $p(U) = \sum_V p(U,V) = \sum_V p(U\cond V)\,p(V)$ for discrete
$V$ respectively.
%
The uninteresting variables $V$ in the present case are the number of
packets $N$, the set of individual luminosities $\bmL$ of the $N$ packets
whose likelihood is modelled by a distribution determined by
likelihood parameters $\bmphi$,
\begin{align}
  \lleq{slk}
  p(\bmL\cond N,\bmphi)
  &= \prod_{j=1}^{N} p(L_j\cond \bmphi)
\end{align}
and the parameter $\lambda$ governing the likelihood for the number of
packets, $p(N\cond\lambda)$. Multiple use of the marginalisation rule
on $p(Q\cond n,\bml)$ results in
\begin{align}
  \lleq{sll}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;\int_0^\infty \rmdx{\lambda}\;
    p(N\cond\lambda)\;
    p(\lambda\cond n)
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    \int \rmdx{\bmphi}\;
    p(\bmL\cond N,\bmphi)\,
    p(\bmphi\cond n,\bml),
\end{align}
where we have omitted from the probabilities all quantities to the
right of the conditional line which are irrelevant to the variable at
hand. For example, Eq.~(\ref{sdf}) shows that the total bin luminosity
$Q$ is fully determined by $N$ and $\bmL$ and so we can shorten
$p(Q\cond N,\bmL,\lambda,\bmphi,n,\bml)$ to $p(Q\cond N,\bmL)$. %
Likewise the probability for $\bmL$ is fully determined by $N$ and the
parameters $\bmphi$ and does not depend on the data $n,\bml$.  The
bounds of $\int \rmdx{\bmL}$ and $\int \rmdx{\bmphi}$ are determined by the
specific model and data for which $p(Q\cond n,\bml)$ is to be
calculated.
%
We note that a part of Eq.~(\ref{sll}) can be seen as a convolution
since $p(Q\cond N,\bmL)$ can be written as a Dirac delta function,
\begin{align}
  \lleq{slm}
  \int \rmdx{\bmL}\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \int \rmdx{\bmL}\;\delta(Q - \textstyle\sum_{j=1}^N L_j)\; \prod_{j=1}^N p(L_j\cond \bmphi).
\end{align}
Eq.~(\ref{sll}) is exact, and in principle all that remains is to make
specific choices for the various probabilities which are judged best
to represent the problem at hand, and thence to carry out the sum and
multiple integrals. In practice, of course, such calculations may be
time-consuming, and there will often be more than one choice for the
functional form of each probability. And those choices affect the
computational effort which may be critical if \refeq{slm} has to be
repeatedly evaluatated for thousands of bins.

\section{Example: modelling single-packet luminosity distributions}
\label{sec:example}

In this section, we show by example how information on the
single-packet luminosity distributions can be utilised to find exact
expressions for Eq.~(\ref{sll}). Our choices are motivated by our
original application \tardis{} but we emphasize that much of the
analysis below, especially \refsec{multiplicity}, is generally
applicable.

\subsection{Multiplicity prediction} \label{sec:multiplicity}

First, the parameter $\lambda$ is eliminated as follows. The number of
packets $N$ is Poisson-distributed,
\begin{align}
  \lleq{pmb}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^{N}}{N!}
  \quad N = 0,1,\ldots,\infty.
\end{align}
The observed $n$ is naturally presumed to have arisen from the same
Poisson distribution with the same $\lambda$.  The posterior for
$\lambda$ is found from the data $n$ via Bayes' Theorem,
\begin{align}
  \lleq{pmc}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
We use an improper prior
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a},
\end{align}
meaning that $\int_0^\infty \rmdx{\lambda}\,p(\lambda)$ is known only up to
a constant $c$. An improper prior is not
a problem because $c$ cancels out when calculating the posterior,
provided that $n \ge 1$; i.e., at least one packet is observed in the
bin.\footnote{For a uniform prior, the case $a{=}0$, we require of
  course a parameter $\lambda_{\rm max}$. As long as
  $\lambda_{\rm max}$ is large enough, it would also cancel.}, The
posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n)
  &= \GammaDist(\lambda \cond \alpha = n{-}a{+}1 \,,\, \beta = 1)\,,
\end{align}
where the $\GammaDist$ distribution with shape parameter $\alpha > 0$ and
rate parameter $\beta > 0$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(x\cond \alpha, \beta)
  &= \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\,,
  \quad 0 \leq x < \infty.
\end{align}
The evidence is
$p(n) =  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda) = c\,(n-a)!/n!$
% \begin{align}
%   \lleq{pmg}
%   p(n)
%   &=  \int_0^\infty \rmdx{\lambda}\, p(n\cond\lambda)\,p(\lambda)
%     \ =\ c\frac{(n-a)!}{n!} \,, \\
% \end{align}
and inserting Eqs.~(\ref{pmb}) and (\ref{pmc}), the prediction for $N$
given $n$ is a negative binomial distribution
\begin{align}
  \lleq{pmh}
  p(N\cond n) %
  &= \int_0^\infty \rmdx{\lambda}\, p(N\cond\lambda)\,p(\lambda\cond n)
  \ =\ \binom{N{+}n{-}a}{N} \left(\frac{1}{2}\right)^{N+n-a+1}\, .
    %\ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
\end{align}
A hyperparameter choice $a=0$ yields the uniform prior, $a=1/2$ gives
the Jeffreys and reference prior, and $a=1$ is the
transformation-group result advocated by
Jaynes~\cite[Ch. 12]{jaynes2003probability}. The uniform prior has no
solid information-theoretic foundation in this case reducing the
choice to $a \in \{1/2, 1\}$. We examine the consequence for the
extreme case that no packet is observed: $P(N=0 \cond n=0) =
2^{a-1}$. For $a=1$, any $N>0$ is excluded which is undesirable unless
we accept that repeating the simulation will never yield a packet in
the current bin. Hence we are led to Jeffreys prior, $a=1/2$.  With
\refeq{pmh}, \refeq{sll} simplifies to
\begin{align}
  \lleq{gmh}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    \int \rmdx{\bmphi}\;
    p(\bmL\cond N,\bmphi)\,
    p(\bmphi\cond n,\bml).
\end{align}

\subsection{Single-packet description} \label{sec:single-packet}

Next, we consider detailed modelling of $p(\bmL\cond N,\bmphi)$ which
was written in Eq.~(\ref{slk}) as a product single-packet
distributions $p(L_j\cond \bmphi)$, meaning that single packets are
assumed to be mutually independent and to follow the same distribution
independent of $N$.

Which specific probability function to choose for $p(L_j\cond\bmphi)$
depends, of course, on the specific simulation data. In this section,
we choose it to be the same Gamma distribution for all $L_j$,
with parameters $\bmphi = (\alpha,\beta)$,
\begin{align}
  \lleq{spc}
  p(L_j\cond \bmphi) &\sim \GammaDist(L_j\cond \alpha,\beta)
\end{align}
which has the nice property that the convolution (\ref{slm}) results
in yet another Gamma distribution,
\begin{align}
  \lleq{spd}
  \int \rmdx{\bmL}\; p(Q\cond N,\bmL)\; p(\bmL\cond N,\bmphi)
  &= \GammaDist(Q\cond N\alpha, \beta),
\end{align}
and so Eq.~(\ref{gmh}) becomes in this example
\begin{align}
  \lleq{spe}
  p(Q\cond n,\bml)
  &= \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    p(\alpha,\beta\cond n,\bml).
\end{align}
The posterior for the parameters is calculated by yet another
application of Bayes' Theorem,
\begin{align}
  \lleq{spf}
  p(\alpha,\beta\cond n,\bml)
  &= \frac{p(n,\bml\cond \alpha,\beta)\,p(\alpha,\beta)}{p(n,\bml)}
\end{align}
where the data $(n,\bml)$ must naturally follow the same
Gamma likelihood as $(N,\bmL)$, i.e.\
\begin{align}
  \lleq{spg}
  p(n,\bml\cond \alpha,\beta)
  & = \prod_{j=1}^n \GammaDist(\ell_j\cond\alpha,\beta)
    \ =\ \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n}
\end{align}
with the data summarized by the sufficient statistics $n$,
$q \equiv \sum_{j=1}^n \ell_j$, and $r \equiv \prod_{j=1}^n \ell_j$.

\fred{Here and above, $n$ does not have to be the same in
  $P(N \cond n)$ and $p(n,\bml\cond \alpha,\beta)$. We can use packets
  from other bins if we assume $\alpha, \beta$ are the same in those bins.}

The only missing piece to explicitly evaluate \refeq{spf} is the prior
$p(\alpha,\beta)$.
% To gain some insight, consider the standard $\GammaDist$ model with
% $\Kalpha = \Kbeta = 0$.
% There is a conjugate prior but
% unfortunately the normalization constant cannot be evaluated in closed
% form, so it is thus of no use for us.
% How about a noninformative prior?
The authors of \cite{moala2013bayesian} present an overview of
possible choices indicating that their \emph{modified MDIP} prior
performed best in the few examples considered there. We do not follow
their recommendation because their modification is not well motivated
and it seems to accidentally perform better because of the particular
examples considered. The other conclusion they have is that the
Jeffreys, reference, and other noninformative priors with similarly
solid motivation have comparable performance. They all exhibit the
standard form for a scale or rate parameter $\propto 1/\beta$ and only
differ in the $\alpha$ dependence. We disregard the reference prior as
it forces us to identify either $\alpha$ or $\beta$ as \emph{the}
nuisance parameter but we consider \emph{both} as nuisance
parameters. Hence we select the Jeffreys prior
\begin{align}
  \lleq{sph}
  p(\alpha,\beta)
  &=  C \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta} \, .
  % \frac{1}{(\alpha_{\rm max}{-}1)\beta_{\rm max}}
  % \qquad 1 < \alpha < \alpha_{\rm max},\; 0 < \beta < \beta_{\rm max}.
\end{align}
% since $\alpha_{\rm max}$ and $\beta_{\rm max}$ cancel in (\ref{spf}).
With these choices, the posterior normalization is
\begin{align}
  \lleq{spj}
  p(n, \bml) &= C \int_0^{\alpha_{\rm max}} d\alpha \int_0^{\beta_{\rm max}} d\beta\;
  \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n},
\end{align}
where the upper integration limits
$(\alpha_{\rm max}, \beta_{\rm max})$ should formally extend to
$\infty$ but in practice have to be chosen by trial and error to be
large enough to contain the relevant contribution to the integral.
The prior's constant of proportionality $C$ cancels in the final
result and we may set $C\equiv 1$ to obtain
\begin{align}
  \lleq{spk}
  p(Q\cond n,\bml)
  &= \frac{1}{p(n, \bml)}
    \sum_{N=0}^\infty \;
    p(N\cond n)\;
    \int \rmdx{\alpha}\,\rmdx{\beta}\;
    \GammaDist(Q\cond N\alpha, \beta)\;
    \frac{\beta^{n\alpha} r^{\alpha-1}\,e^{-\beta q}}{[\Gamma(\alpha)]^n} \frac{\sqrt{\alpha \Psi'(\alpha)-1}}{\beta}.
\end{align}

\subsection{Preprocessing \tardis{} output } \label{sec:tardis}

\hans{H: This subsection needs much more work. I have just copied some
  relevant text in here which may be useful. Much of it should be
  removed. Some connection to TARDIS should preferably be kept in the
  paper, in my view, eg the definition of $X = 1 - L/\ell_{\rm max}$.}

The \tardis{} packages simulates supernovae with %
\fred{Wolfgang's favorite description of his package}. %
Here we focus on the real packets as opposed to the virtual packets
output by tardis.

A histogram plot of the luminosities from a \tardis{} run indicates a
strongly skewed distribution with its single mode close to the maximum
luminosity $\lmax$, while there is a long tail for small $\ell$. These
features can be accomodated by a $\GammaDist$ distribution if we
transform the luminosities to
\begin{align}
  \lleq{gmf}
  x_j \equiv 1- \frac{\ell_j}{\lmax} > 0
\end{align}
and similarly $X_j = 1 - (L_j/\lmax)$ and $\Lumtot \to X$. We assume
the total number of samples across all bins is large, hence the effect
of considering $\lmax$ a constant independent of the data is
negligible. We checked that leaving $\lmax$ as a free parameter in a
fit gave equivalent results.
%
Numerically, we take the largest luminosity of all \tardis{} packets
across the entire frequency range and increase it slightly to avoid
numerical difficulties at the end point
\begin{align}
  \lleq{gme}
  \lmax \equiv (1+\epsilon) \times \max_i \ell_i,
\end{align}
where $ \epsilon=10^{-6}$.

We infer $p(X \cond n,\bml)$ and only in the end transform back to a
probability density in $\Lumtot$ via
\begin{align}
  \lleq{gmg}
  p(\Lumtot  \cond n, \bml)
  &= p(X  \cond n, \bmx) \frac{1}{\lmax} \,.
\end{align}
To simplify the numerics, we also rescale the frequencies by a
constant from $\mathcal{O}(10^{15})$ to $\mathcal{O}(1)$.  %

A major numerical simplification arises because of the $\Lumtot \to X$
transformation. Rewriting the \refeq{sll} formally as
\begin{align}
  \lleq{gmhb}
  p(Q  \cond n,\bml)
  & = \sum_{N} p(N \cond n)\; p(Q \cond N, n, \bml),
\end{align}
we see that the right-hand side is a mixture density. In our
application, the distribution of a single-packet luminosity is narrow
in the sense that the standard deviation relative to the mean is at
the percent level. This implies that $p(Q \cond N, n, \bml)$ has one
mode from each $N$ and all modes are clearly separated. The
transformation $\Lumtot \to X$ effects a strong overlap between the
distributions $p(X \cond N, n, \bmx)$ with similar $N$, so
$p(X \cond n, \bmx)$ can be well approximated by a unimodal
distribution.

We chose the Gamma distribution because it is \emph{stable}; i.e.,
the sum of $N$ Gamma variates is again a Gamma variate
\begin{align}
  \lleq{gmb}
    X_j \sim \GammaDist(\alpha, \beta) \Rightarrow X \equiv \sum_{j=1}^N X_j \sim \GammaDist(N \alpha, \beta).
\end{align}
Assuming $\alpha = \alpha(\nu \cond \bmphi)$ and $\beta = \beta(\nu
\cond \bmphi)$ are functions of $\nu$ that depend on parameters
$\bmphi$, we set
\begin{align}
  \lleq{gmib}
  p(\bmX \cond \bmphi, N, \nu) &= \prod_{j=1}^N \GammaDist(X_j \cond \alpha, \beta ),\\
  p(\bmx \cond \bmphi, \npack, \bmnu) &= \prod_{i=1}^{\npack} \GammaDist(x_i \cond \alpha_i, \beta_i).
\end{align}
where $\alpha_i \equiv \alpha(\nu_i \cond \bmphi), \beta_i \equiv
\beta(\nu_i \cond \bmphi)$.

\subsection{Linear model}

\fred{Parts may still be relevant, especially if we put a plot of an
  entire spectrum in the paper.}

For simple functions $\alpha(\nu \cond \bmphi)$, one can find the
minimum value of $\alpha$ over the entire range of $\nu$
analytically. If this is too complicated, we propose to require
$\alpha(\nu_i \cond \bmphi) > 1$ only at each packet frequency
$\nu_i$. In the evaluation of $p(\bmx \cond \bmphi, \npack, \bmnu)$,
one has to loop over all packets and compute $\alpha$ and $\beta$
anyways, so the constraint imposes essentially no computational
overhead.

In the absence of a physical model to determine
$\alpha(\nu \cond \bmphi)$ and $\beta(\nu \cond \bmphi)$, we propose
to start with a simple polynomial ansatz
$\alpha(\nu \cond \bmphi) = \sum_{k=0}^\Kalpha \alpha_k \nu^k,
\beta(\nu \cond \bmphi) = \sum_{k=0}^\Kbeta \beta_k \nu^k$ such that
\begin{align}
  \lleq{gmp}
  \bmphi = (\alpha_0, \dots \alpha_\Kalpha , \beta_0, \dots \beta_\Kbeta)
\end{align}
The usual trade off between parsimony and the ability to accurately
model the frequency dependence determines the orders of the
polynomials, $\Kalpha$ and $\Kbeta$. This is outside the scope of this
paper. Here we choose $\Kalpha = 3$ and $\Kbeta = 1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Large packet number approximations} \label{sec:asymptotic}

In practice, the situation may arise that many packets fall into a
bin. One can then avoid modeling the luminosity distribution
explicitly as done in \refsec{single-packet} as it becomes irrelevant
in the limit $n \to \infty$, essentially due to the central limit
theorem. As a bonus, the numerical effort to evaluate the basic
\refeq{sll} drastically simplifies.

\subsection{Derivation of $p(Q\,|\,n,\boldmath{\ell})$ for large $n$}

We return to the general formulation of \refeq{sll} and re-arrange its
factors into
\begin{align}
  \lleq{asc}
  p(Q\cond n,\bml)
  &= \int \rmdx{\lambda}\; \rmdx{\bmphi}\;
    p(\lambda\cond n)\;
    p(\bmphi\cond n,\bml)\;
    p(Q\cond \lambda,\bmphi), \\
  \lleq{asd}
  p(Q\cond \lambda,\bmphi)
  &= \sum_N
    p(N\cond \lambda)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    p(\bmL\cond N,\bmphi),
\end{align}
with $p(\bmL\cond N,\bmphi) = \prod_{j=1}^N p(L_j\cond\bmphi)$ as
before. Again all quantities are related to a single bin in frequency.
Writing the moments and variance of $L$ as\footnote{The $k$-th moment
  of variable $x$ is defined as the expectation value $\mu_k =
  \expect{x^k} = \int \rmdx{x}\,x^k\,p(x),\; k = 0,1,2,\ldots$.} %
\begin{align}
  \lleq{ram}
  \mu =  \mu_1 &\equiv \expect{\Lum}, \qquad
  \mu_2 \equiv \expect{\Lum^2}, \\
  \sigma^2 &\equiv \variance{\Lum}
  = \expect{\Lum^2} - \expect{\Lum}^2,
\end{align}
it is easy to show that, for the Poissonian compound process
(\ref{asd}), the first moment and variance of $Q$ are related to those
of $L$ by
\begin{align}
  \lleq{raa}
  \expect{Q} = \lambda \mu, \qquad
  \variance{Q} = \lambda \mu_2 = \lambda (\mu^2 + \sigma^2) \,.
\end{align}
for any probability $p(L\cond\bmphi)$.
%
Higher-order quantities such as the skewness and kurtosis\footnote{
  The variance is the second so-called \textit{cumulant}. The relation
  for all orders is that the $k$-th order cumulant $\kappa_q(Q)$ is
  equal to $\lambda$ times the $k$-th order moment of $L$,
  $\mu_k$. The skewness of $p(Q\cond \lambda,\bmphi)$ is therefore
  $\kappa_3(Q)/\kappa_2(Q)^{3/2} = \lambda \mu_3/(\lambda \mu_2)^{3/2}
  = \lambda^{-1/2} \mu_3/\mu_2^{3/2}$.}  %
are suppressed by powers of $\lambda^{-1/2}$.
%%%%%%%%%%%
\hans{H comment (not for text): The magnitude of the correction to
  Eq.~(\ref{rab}) due to $\kappa_3$ and higher orders is unclear.}
%%%%%%%%%%%
From Theorem 4.3.1 of Ref.~\cite{bening2002generalized}, we conclude
that the asymptotic distribution of $\Lumtot$ when
$\lambda \to \infty$, given the first and second moment of $\Lum$, is
a Gaussian or normal distribution,
\begin{align}
  \lleq{rab}
  p(\Lumtot \cond \lambda,\bmphi)
  &= \GaussianDist \left( \Lumtot \cond \expect{Q}, \variance{Q} \right) %
    = \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 {+} \sigma^2) \right).
\end{align}
This drastic simplification eliminates both the sum over $N$ and the
integral over $\bmL$ in \refeq{asd}.

It remains to find the integrals over $\lambda$ and $\bmphi$ in
\refeq{asc}. As before, the posterior for $\lambda$ is the Gamma
distribution of \refeq{pms}. %
%
To determine $p(\bmphi\cond n,\bml)$, we note that the result
(\ref{rab}) also simplifies the calculation of the parameter posterior
$p(\bmphi\cond n,\bml)$ since, independently of the actual functional
model for $p(L\cond\bmphi)$ the only relevant information is contained
in the first moment $\mu$ and variance $\sigma^2$.  As set out in
Jaynes~\cite[Ch. 7.10]{jaynes2003probability}, knowledge of $\mu$ and
$\sigma^2$ implies that we can assign a Gaussian distribution to
$p(L\cond\bmphi)$
\begin{align}
  \lleq{rae}
  p(\Lum \cond \bmphi) =
  p(\Lum \cond \mu, \sigma^2 ) &= \GaussianDist(\Lum \cond \mu, \sigma^2)
\end{align}
even though the actual data histograms may not be Gaussian. The reason
is that a Gaussian represents our state of knowledge \emph{before}
data $\bmy$ is observed. With this assignment,
$\bmphi =(\mu,\sigma^2)$ and the posterior becomes, using Bayes'
Theorem,
\begin{align}
  \lleq{rac}
  p(\mu,\sigma^2\cond n,\bml)
  &= \frac{
    \GaussianDist(\bml\cond\mu,\sigma^2)\;
    p(\mu\cond n)\;
    p(\sigma^2\cond n)}
  {p(\bml\cond n)}
\end{align}
In keeping with prior information that luminosities are always
positive, we assign an improper uniform prior
$p(\mu\cond\mu_{\rm max}) = 1/\mu_{\rm max}$ independent of $n$; as
always, the posterior is independent of
$\mu_{\rm max}$.
For $\sigma^2$, we use an $n$-independent Inverse Gamma
distribution prior with hyperparameters $(a_0,b_0)$ which reads
\begin{align}
  \lleq{rad}
  p(\sigma^2\cond a_0,b_0)
  = \InvGammaDist(\sigma^2\cond a_0,b_0)
  &= \frac{b_0^{a_0}}{\Gamma(a_0)} (\sigma^2)^{-a_0-1} \exp(-b_0/\sigma^2)
\end{align}
and is ``conjugate'' to the normal distribution likelihood of
\refeq{rae} in that the posterior for $(\mu,\sigma^2)$ is also a
Gaussian in $\mu$ and an Inverse Gamma in $\sigma^2$, but with
parameters updated by the data,
\begin{align}
  \lleq{rbd}
  p(\mu,\sigma^2\cond n,\bml,a_0,b_0)
  &= \GaussianDist(\mu\cond \mu_n,\sigma_n^2)
  \cdot
  \InvGammaDist(\sigma^2\cond a_n,b_n) \\
  \mu_n &= \langle\ell\rangle\\
  \sigma_n^2 &= \sigma^2/n \\
  a_n &= a_0 + \tfrac{1}{2}n \\
  b_n &= b_0 + \tfrac{1}{2} n\left( \langle\ell^2\rangle - \langle\ell\rangle^2\right)
\end{align}
with data sample moments $\langle\lum^k\rangle$ defined as
$(1/n) \sum_{j=1}^n \lum_j^k$. Since we assume $n \gg 1$, we may
choose $a_0 = b_0 = 0$ resulting in a noninformative prior on $\sigma^2$.
%
Inserting Eqs.~(\ref{pms}) and (\ref{rbd}) into \refeq{asc}, we obtain
\begin{align}
  \lleq{ral}
  p(\Lumtot \cond n,\bml)
  &= \int \rmdx{\lambda}\; \rmdx{\mu}\; \rmdx{\sigma^2}\;
  \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\sigma^2{+}\mu^2) \right)
  \times\\
  &\quad\times
  \GammaDist(\lambda \cond n{-}a{+}1 , 1) \cdot
  \GaussianDist(\mu\cond \mu_n,\sigma_n^2) \cdot
  \InvGammaDist(\sigma^2\cond a_n,b_n) \;
  \nonumber
\end{align}

\subsection{Numerical implementation}\label{sec:asympt-numeric}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{asymptotic}
  \caption{Numerical prediction of $\Lumtot$ with the sum over $N$
    from \refeq{spk} or the asymptotic expression from \refeq{ral} for $n$
    samples from $\GammaDist(\alpha=1.5, \beta=60)$}
\label{fig:asymptotic}
\end{figure}

In \reffig{asymptotic}, we compare the asymptotic expression
\refeq{ral} with the ``exact'' expression \refeq{spk} where the
integrals over $\bmphi$ are done numerically to a requested relative
accuracy of $10^{-5}$ through the Cubature
package\cite{cubature,genz1980remarks,berntsen1991adaptive}.

For large $n$ where the asymptotic expression \refeq{ral} is valid,
asymptotic normality justifies the use of the Laplace or saddle-point
approximation~\cite[Ch. 27]{mackay2003information}. We implemented
\fredmargin{\href{https://github.com/tardis-sn/XXX}{github}}
the
above formulae in the julia language~\cite{julia14} and used Newton's
method from the \texttt{Optim.jl} package in combination with
automatic differentiation~\cite{RevelsLubinPapamarkou2016} to optimize
the respective integrand and to compute the Hessian at the mode as
required for the Laplace method. As initial values for optimization,
we slightly offset the sample estimates
\begin{align}
  \lleq{rat}
  \qty(\lambda, \mu, \sigma^2) = 1.0005 (n, \langle\lum\rangle, \langle\lum^2\rangle{-}\langle\lum\rangle^2)
\end{align}
to avoid issues if the optimizer cannot improve on the initial guess
and found convergence within $\order{10}$ steps.

As a test case, we generated
$\ell_i \sim \GammaDist(\alpha=1.5, \beta=60), i=1 \dots n$, and
computed $P(\Lumtot \cond n, \bml)$ on a grid of $\Lumtot$ values and
checked if the probability integrates to one using Simpson's
rule. With the exact expression, we obtain 0.997 ($n=10$) and 0.996
($n=80$), and 0.698 and 0.953 for the asymptotic expression. For
$n=10$, the large discrepancy arises mostly because the positivity
constraint $P(Q \leq 0 \cond n, \bml) = 0$ is not respected in
\refeq{ral} since the Gaussian approximation \refeq{rae} allows
negative $\lambda$ and $\Lumtot$, whereas the results \refeq{spk}
based on the $\GammaDist$ model imply $\lambda \geq 0$ and $Q \geq 0$.
%
\fred{Update values after Jeffreys implementing prior}

For that reason, we discourage the asymptotic approximation for small
$n$ but we note that in general, the lack of numerical normalization
may not be a concern. For example, if \refeq{ral} is used as the
likelihood for a single bin in Bayesian parameter inference, the
normalization or evidence is
irrelevant. Following~\cite{tierney_accurate_1986}, we renormalized
the results to integrate to one for ease of comparison in
\reffig{asymptotic}.  We verified that for $n>80$, the two methods
agree even better than shown there and that the normalization of the
asymptotic result approaches one. Integrating the asymptotic
expression with cubature instead of Laplace gave very similar results
but turns out to be less reliable and requires many more iterations.

The main advantage of the approximation is computational efficiency. A
single evaluation of $P(Q \cond n, \bml)$ with the Laplace methods
requires only about 100 calls to the integrand of \refeq{ral}
regardless of the value of $n$. In contrast, the numerical integration
of a single term for fixed $N$ in \refeq{spk} requires 1000s of
function evaluations, and the number of terms grows like $\sqrt{n}$
for a given accuracy of $P(Q \cond n, \bml)$.

\section{Conclusion} \label{sec:conclusion}

TODO

\section{Appendix} \label{sec:appendix}

\subsection{Moments of the luminosity prediction}\label{sec:predict-moments}
\hans{H: I would agree to remove this section; I do not see value added.}

While the integral in \refeq{ral} cannot be integrated analytically,
the integrand is unimodal and strongly peaked, so that the Laplace or
saddle-point approximation~\cite[Ch. 27]{mackay2003information} can be
applied. If only the value $p(\Lumtot{=}\lumtot \cond n,m_1,m_2)$ at the
total data luminosity $q$ is sought, then one can approximate it
directly. In addition, we can compute the moments of $\Lumtot$ through
the \emph{moment generating function} \cite{stuart1994kendall}
\begin{align}
  \lleq{rap}
  M(t \cond \bmy) = \int_0^{\infty} \rmdx{\Lumtot} p(\Lumtot \cond \bmy) e^{-t \Lumtot} \,
\end{align}
from which the moments are accessible by differentiation
\begin{align}
  \lleq{raq}
  \expect{\Lumtot^k} = (-1)^k \eval{\dv[k]{M(t \cond \bmy)}{t}}_{t=0}
\end{align}
Assuming all involved integrals are absolutely convergent, we can
exchange the order of integration and find
\begin{align}
  \lleq{rar}
  M(t \cond \bmy) &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond \bmy) p( \mu, \sigma^2 \cond \bmy)  f(t \cond \lambda, \mu, \sigma^2),
  \intertext{where }
  f\left(t \cond  \lambda, \mu, \sigma^2\right) &\equiv \int \rmdx{\Lumtot} \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 + \sigma^2) \right) e^{-t \Lumtot}\\
  &= \frac{1}{2} \exp \qty(t/2 \lambda (t (\mu^2 + \sigma^2)- 2\mu)) \erfc \qty( \frac{\sqrt{\lambda}(t \qty(\mu^2 + \sigma^2) - \mu)}{\sqrt{2 \qty(\mu^2 + \sigma^2)}}).
\end{align}
This implies that the mean and variance of $\Lumtot$ require an integral over the three parameters
\begin{align}
  \lleq{ras}
  \expect{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty(-\eval{\dv{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0})\\
  \variance{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty(\eval{\dv[2]{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0}) - \expect{\Lumtot}^2
\end{align}
We implemented the above formulae in the julia language~\cite{julia14}
and used Newton's method from the \texttt{Optim.jl} package in
combination with automatic
differentiation~\cite{RevelsLubinPapamarkou2016} to optimize the
respective integrand and compute the Hessian at the mode. As initial
values, we slightly offset the sample estimates to avoid issues if the
optimizer cannot improve on the initial guess and found convergence
within $\order{10}$ steps:
\begin{align}
  \lleq{rat}
  \qty(\lambda, \mu, \sigma^2) = 1.0005 (n, \langle\lum\rangle, \langle\lum^2\rangle{-}\langle\lum\rangle^2).
\end{align}
\fred{At least for us, I should do a comparison plot to evaluate $p(\Lumtot \cond \bmy)$ with both methods } %
\fred{We need a plot, the one to compare predictions and 250 real replicas or 100 virtual replicas?} %
\fred{Advantage of this approach: no need to store all packets anymore, sum and sum of squares can be accumulated online $\Rightarrow$ reduced computation and storage requirements}

\section{Delendum}

\subsection{Numerical implementation} \label{sec:numerical}

\fred{Editing will be done by Fred}

In our numerical implementation, the constants
$(\alpha_{\rm max}, \beta_{\rm max})$ are chosen by trial and error to
be large enough to contain the relevant region of
$p(\ell\cond \alpha,\beta)$.
%
To perform the integrals over $(\alpha,\beta)$ to find $C$ and the
Gamma-weighted numerator in (\ref{spe}), we used the fact that the
posterior $p(\alpha,\beta \cond n,\bml)$ is approximately Gaussian in
both $\alpha$ and $\beta$ for large $n$, so we can apply Laplace's
method for both. This requires the Hessian at the mode and the
gradient helps in locating the mode. The relevant analytic expressions
are summarized in \refsec{appendix}.

Regarding the sum over $N$, we start at
$N = \lfloor n-a+1 \rfloor = \arg \max_N p(N|n)$ %
\fred{Actually we should say $p(N \cond n, a)$} %
and continue to larger and smaller values of $N$ until the
contribution to $p(Q \cond n,\bml)$ becomes negligible. In case the
relevant contribution is from large $N$ and $n$, we interpret $N$ as a
continuous variable, replace factorials by $\Gamma$ function in
\refeq{spj}, and apply Laplace's method to integrate over $N$ also.

 \subsection{Poisson-gamma convolution}\label{sec:poisson-gamma}

\hans{H: This subsection is temporary. It is motivated by the question
  whether the use of a gaussian for the single-packet luminosity
  $p(L_j\cond\bmphi)$ in (\ref{rae}) is necessary and whether making
  it a gaussian is a source of bias since $L_j$ cannot be
  negative. The original choice of a Gamma distribution as in
  (\ref{spc}) is more appropriate and could eliminate one source of
  systematic bias. Below, I have derived a closed form for
  $p(Q\cond \lambda,\bmphi)$ using the gamma-distributed
  $p(L_j\cond \alpha,\beta)$ of (\ref{spc}). The final result
  (\ref{qfp}) below could be used to replace the gaussian result
  (\ref{rab}) for insertion into (\ref{asd}). Inclusion of the entire
  derivation below is not envisaged; rather, we would include only
  salient issues and equation in the text itself. For the moment, the
  question is what Fred has to say about this version ie whether it
  can be implemented numerically and how it compares to the above
  results.

  The limit of this section to Gaussian can be investigated later if necessary.
}

We start again from (\ref{asd}) and use (\ref{spc})
\begin{align}
  \lleq{asdd}
  p(Q\cond \lambda,\bmphi)
  &= \sum_N
    p(N\cond \lambda)\;
    \int \rmdx{\bmL}\;
    p(Q\cond N,\bmL)\;
    p(\bmL\cond N,\bmphi)\\
  \lleq{spcd}
  p(L_j\cond \alpha,\beta)
  &= \frac{\beta^\alpha L_j^{\alpha-1}e^{-\beta L_j}}{\Gamma(\alpha)}.
\end{align}
The moment generating function of $p(L_j\cond \alpha,\beta)$ is
\begin{align}
  M[t\cond p(L_j\cond \alpha,\beta) ]
  &= E[e^{tL_j}] = \int_0^\infty dL_j e^{tL_j}\,p(L_j\cond \alpha,\beta)
    = (1 - t/\beta)^{-\alpha}
\end{align}
The convolution of $N$ Gamma distributions yields the moment generating
function
\begin{align}
  M[t\cond p(Q\cond \alpha,\beta) ]
  &= (1-t/\beta)^{-N\alpha}
\end{align}
and so we obtain the well-known result that the convolved distribution
is also a Gamma distribution with shape parameter $N\alpha$,
\begin{align}
  p(Q\cond \alpha,\beta)
  &= \frac{\beta^{N\alpha} Q^{N\alpha-1}e^{-\beta Q}}{\Gamma(N\alpha)}.
\end{align}
Inserting this into the $N$-sum, we have
\begin{equation}
  \lleq{gms}
  \boxed{
  p(Q\cond \lambda,\alpha,\beta)
  = \frac{e^{-\lambda - \beta Q}}{Q}
    \sum_{N=0}^\infty
    \frac{[\lambda \beta^\alpha Q^\alpha]^N}{N!\;\Gamma(N\alpha)}.
    }
\end{equation}
This series itself should converge quickly because of the
$N!\;\Gamma(N\alpha)$ in the denominator. It is possible to find an
analytic solution for some special cases of $\alpha$.
\begin{itemize}
\item For $\alpha=1$ we have
  \begin{align}
    p(Q\cond \lambda,\alpha{=}1,\beta)
    &= \frac{e^{-\lambda - \beta Q}}{Q}
      \sum_{N=0}^\infty
      \frac{[\lambda \beta Q]^N}{N!\;(N-1)!}.
  \end{align}
  Comparing the above to the series definition of the modified Bessel function,
  \begin{align}
    I_1(z) &= \sum_{k=0}^\infty \frac{(z/2)^{2k+1}}{k!\;(k+1)!},
  \end{align}
  we can identify $z = 2\sqrt{\lambda\beta Q}$ to get
  \begin{equation}
    \boxed{
      p(Q\cond \lambda,\alpha{=}1,\beta)
      = \sqrt{\frac{\lambda\beta}{Q}}  e^{-\lambda - \beta Q}
      \;I_1\!\left(2\sqrt{\lambda\beta Q}\right)
      }
  \end{equation}
  as derived by Fisher and Cornish (1960). We often prefer to use the
  corresponding hypergeometric function formulation, in terms of which
  \begin{align}
    p(Q\cond \lambda,\alpha{=}1,\beta)
    &= \lambda\beta\; e^{-\lambda - \beta Q}
      \;{}_0F_1(2; \lambda\beta Q).
  \end{align}

\item For integer $\alpha = 2,3,\ldots$, we rely on a paper by Withers and
  Nadarajah (2011) from which we get an expression for the (p,q) hypergeometric
  function, defined as
  \begin{align}
    {}_pF_q\left( a_1,a_2,\ldots, a_p\;;\; b_1,b_2,\ldots,b_q\;;\; z \right)
    &= \sum_{k=0}^\infty \frac{z^k}{k!} \;
      \frac{(a_1)_k(a_2)_k\cdots (a_p)_k}{(b_1)_k(b_2)_k\cdots (b_q)_k}
  \end{align}
  in terms of the Pochhammer symbol $(a)_k = \Gamma(a+k)/\Gamma(a)$. Withers
  and Nadarajah found that (and I checked)
  \begin{align}
    (N\alpha)!
    &= (\alpha^\alpha)^N
      \left(\frac{1}{\alpha}\right)_k
      \left(\frac{2}{\alpha}\right)_k
      \left(\frac{\alpha{-}1}{\alpha}\right)_k
      \left(\frac{\alpha}{\alpha}\right)_k
  \end{align}
  which, when inserted into the appropriate hypergeometric function
  ${}_0F_\alpha$, yields
  \begin{align}
    \lleq{qpfa}
    {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
    \frac{1}{\alpha},
    \frac{2}{\alpha},
    \ldots,
    \frac{\alpha{-}1}{\alpha},
    \frac{\alpha}{\alpha}
    \;;\; z \right]
    &= \sum_{k=0}^\infty \frac{z^k}{k!} \;\frac{(\alpha^\alpha)^k}{(k\alpha)!}
  \end{align}
  and with the general relation (see DLMF website)
  \begin{align}
    z\frac{d}{dz} {}_0F_p(\;;\;b_1,b_2,\ldots,b_q\;;\;z)
    &= \frac{z}{b_1b_2\cdots b_q}
      {}_0F_p(\;;\;b_1{+}1,b_2{+}1,\ldots,b_q{+}1\;;\;z)
  \end{align}
  we obtain, with the assignment
  \begin{align}
    z &= \lambda (\beta Q / \alpha)^\alpha \\
    \lleq{qpfe}
    p(Q\cond \lambda,\alpha,\beta)
      &= \frac{\alpha e^{-\lambda - \beta Q} }{Q}
        z\frac{d}{dz} \;
        {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
        \frac{1}{\alpha},
        \frac{2}{\alpha},
        \ldots,
        \frac{\alpha{-}1}{\alpha},
        \frac{\alpha}{\alpha}
        \;;\; z \right]
        \nonumber\\
      &= \frac{\alpha e^{-\lambda - \beta Q} }{Q}
        \frac{z}{(\alpha-1)!\;\alpha^{\alpha-1}} \;
        {}_0F_\alpha\left[ 0,0,\ldots,0 \;;\;
        1{+}\frac{1}{\alpha},
        1{+}\frac{2}{\alpha},
        \ldots,
        1{+}\frac{\alpha{-}1}{\alpha},
        1{+}\frac{\alpha}{\alpha}
        \;;\; z \right]
        \nonumber\\
  \end{align}
  which on insertion of $z$ and simplification yields for integer $\alpha$
  \begin{equation}
    \lleq{qfp}
    \boxed{
      p(Q\cond \lambda,\alpha,\beta)
      =
    \frac{\lambda\beta^\alpha Q^{\alpha-1} e^{-\lambda - \beta Q} }{(\alpha-1)!}\;
    {}_0F_\alpha\left[ \;;\;
        \frac{\alpha{+}1}{\alpha},
        \frac{\alpha{+}2}{\alpha},
        \ldots,
        \frac{2\alpha{-}1}{\alpha},
        2
        \;;\; \lambda\left(\frac{\beta Q}{\alpha}\right)^\alpha \;\right]
      }
  \end{equation}
  which includes the $I_1$ Bessel function for $\alpha=1$. Mathematica
  has the function ${}_pF_q$ available. Where it is not available, the
  above result should reduce, of course, to the series (\ref{gms}).
\end{itemize}

\bibliographystyle{plain}
\bibliography{references}

\end{document}

% Local Variables:
% compile-command:"rubber --pdf -W refs --synctex draft"
% End: